Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-11 writedown/period.noneq.msm/paper.tex
added conclusion. worrying about the Markovianity of the L(t)

\section{Concluding remarks}

In this paper we present the way of building MSM for a periodically
driven non-equilibrium system.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{fig/flowchart/flowchart.eps}
  \caption{The flowchart of this work.}
  \label{fig:flowchart}
\end{figure}

>>>>

\section{Concluding remarks and discussions}

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{fig/flowchart/flowchart.eps}
  \caption{The flowchart of this work. A lighter color indicates
  a ``more coarse'' approximation to the original NEMD dynamics.}
  \label{fig:flowchart}
\end{figure}

In this paper we methods of building MSM for a periodically driven
non-equilibrium system. The methods are shown, by a numerical example
of the alanine dipeptide under periodically oscillatory EF, to be
valid and successfully captures the long-time behavior of the original
non-equilibrium dynamics. We have provided and verified two equivalent
pathways of building MSM, as is shown by the two branches in
Fig.~\ref{fig:flowchart}.

The the key steps of the first pathway are
summarized as following:
\begin{enumerate}
\item By a fine discretization of the molecular configuration space,
  the original NEMD dynamics is discretized into a (in principle)
  time-continuous \recheck{Markov} jump process with a time-dependent generator $L(t)$.
  If the discretization is good enough, then the discretized
  dynamics is expected to reproduce the original dynamics.
  In principle most of the discretiztion methods for the
  equilibirum case can be generalized to  non-equilibrium without
  substantial difficulty, however,
  we have not discussed which discretiztion method is the best choice,
  since it is out of the scope of the present work.
  \recheck{It should be noticed that
    we assume the jump process is Markovian, althogh in
    general it is not true. However, if (1) the system present time-scale
    seperation; (2) the spatial discretiztion is fine enough; (3)
    we approximately consider a discretizted time
    (can be very fine, and in computer the molecular configurations are
    saved at discretizted time)
    by lag-time $\tau$, so that within the lag-time
    the fast dynamics of the system is fully relaxed; then
    the Markovianity is a good assumption for the discretizted dynamics.
  }
  We show that, by numerical example, a uniform $20\times20$ discretiztion 
  over the dihedral angles $\phi$--$\psi$ space is a good choice for
  alanine dipeptide.
  In the tests, the state independency of computing $L(t)$
  is used as an indicator for
  the quality of the discretiztion.
\item By the Floquet thoerem, the time-inhomogeneous process generated
  by $L(t)$ is temporally discretized in to a time-homogeneous,
  but not necessarily reversible, jump process that has a transition
  matrix denoted by $P$. This temporal discretiztion only  considers
  the state of the system at integral periods, therefore all information
  within one-period is not preserved. This is, however, not a serious
  problem whenever the long-time hehaviors of the system is of interst, and
  their corresponding time-scales are significantly longer than one period.
  The advantage of this discretiztion
  is also obvious: All well-developed theories and computational tools
  for time-homogeneous Markov processes
  can be directly appled. In the
  alanine dipeptide system, we compare
  the the quasi-stationary probability,
  first mean hitting time, and the forward and backward committors
  of the discretized dynamics with those computed directly from the
  NEMD simulation. An almost perfect consistency 
  indicates a good approximation quality of the discretized dynamics.
\item From the disretized dynamics, building the MSM is straightforward
  by using Eq.~\eqref{eq:msm-tmatrix-00} and a set of coresets.
  Numerical results shows that a 3-state MSM can reporduce the leading
  non-trivial eigenvalue with very good accuracy, while the second
  non-trivial eigenvalue with acceptable accuracy. The lower accuracy
  regarding the second non-trivial eigenvalue may be due to the fact that
  the second slowest time-scale is not significantly longer than the
  period. Using this three state MSM, we can also reproduce, with
  almost perfect accuracy, the 
  time-evolution of the metastable state
  pobabilities from equilibirum (no external driven)
  to the non-equilibrium quasi-stationary state.
\end{enumerate}

The right branch in Fig.~\ref{fig:flowchart} presents a
equivalent, and seemingly
much simpler alternative to the left branch, provided the
coresets are known. Neverthless, in pactice, identifying
the coresets is not a trivial task, especially for the
non-equilibrium molecular systems.
Therefore, a good discretization is  usually a must for computing the coreset.
For example, in the present work, we computed the coresets by
finding the roughly constant levels of the hitting time distribution.
This method only works for the discretized dynamics $P$, and its continuous
counterpart is still an open question.

In this paper, we mainly focus on the methodology development for MSM
building in non-equilibrium systems. It is worth noting that the
numerical example itself presenst interesting phenomena: under the
oscillatory EF, the population of the left-handed $\alpha$-helix
significantly increases with respect to the equilibirum population
(see also the discussions in Ref.~\cite{wang2014exploring}).  Along
with building MSM, it is in principle possible to apply the non-equilibrium
linear response theory~\cite{wang2013linear} to build a MSM for the perturbed
system~\cite{schutte2014markov} with respect to original
non-equilibrium system. This makes possible the molecular design under
oscillatory EF, and could be used to answer the question like:
in order to achieve a certain population of  the left-handed $\alpha$-helix,
what kind of EF should be used to drive the system.
In this work, we have not investigated in this direction, however, it will
be treated in the future studies.
